---
title: "Analysis of Fetal Membrane at Midterm and Term, Cell Cell Comunication"
author: "Lukas Moser"
date: "`r format(Sys.time(), '%d %B %Y')`"
format:
  html:
    self-contained: true
---

According to Vignette: 
## Reading the file

```{r, message=FALSE, warning=FALSE, echo=FALSE}
# Load Seurat library
library(Seurat)
library(SingleCellExperiment)
library(dplyr)
library(ggplot2)
library(nichenetr)
library(multinichenetr)

# Read the RDS file
amnion_seurat_object <- readRDS(file = "amnion_seurat_object.rds")

```

```{r}
counts <- GetAssayData(amnion_seurat_object[["RNA"]], slot = "counts")
metadata <- amnion_seurat_object@meta.data

# -------------------------
# 2) Build new Seurat object
# -------------------------
obj <- CreateSeuratObject(
  counts = counts,
  meta.data = metadata,
  assay = "RNA",
  project = "Amnion_Rebuild"
)

# -------------------------
# 3) Preprocessing
# -------------------------
# Normalize
obj <- NormalizeData(obj, normalization.method = "LogNormalize", scale.factor = 1e4)

# Find variable features
obj <- FindVariableFeatures(obj, selection.method = "vst", nfeatures = 2000)

# Scale data
obj <- ScaleData(obj, features = rownames(obj))

# -------------------------
# 4) PCA
# -------------------------
obj <- RunPCA(obj, features = VariableFeatures(obj), verbose = FALSE)

# -------------------------
# 5) Neighbors and Clustering
# -------------------------
obj <- FindNeighbors(obj, dims = 1:20)
obj <- FindClusters(obj, resolution = 0.5)

# -------------------------
# 6) UMAP
# -------------------------
obj <- RunUMAP(obj, dims = 1:20)

amnion_seurat_object <- obj

rm(obj)
rm(counts)

gc()

```


```{r}
amnion_no_graphs <- amnion_seurat_object
amnion_no_graphs@graphs <- list()

all_seurat_object_filtered <- subset(amnion_no_graphs, subset = timepoint %in% c("Week_25", "Week_38"))
all_seurat_object_filtered <- subset(all_seurat_object_filtered, subset = tissue_source %in% c("Peripher"))
all_seurat_object_filtered$samples <- paste(
  all_seurat_object_filtered$timepoint,
  all_seurat_object_filtered$Donor_Nr,
  sep = "_")
Idents(all_seurat_object_filtered) <- "samples"
int_seurat_object <- NormalizeData(all_seurat_object_filtered)
int_seurat_object <- FindVariableFeatures(int_seurat_object, selection.method = "vst", nfeatures = 50000)
int_seurat_object <- ScaleData(int_seurat_object)

```

```{r}
rm(amnion_no_graphs)
rm(amnion_seurat_object)
rm(all_seurat_object_filtered)
gc()
# Access metadata
metadata <- int_seurat_object@meta.data

# Create the group_time column based on the existing group column
metadata$group_time <- metadata$timepoint
# Update the metadata in the Seurat object
int_seurat_object@meta.data <- metadata
gc()
```

```{r}
# Access metadata
metadata <- int_seurat_object@meta.data

# Replace spaces, plus signs, and hyphens with underscores in the cell_type column
metadata$cell_type <- gsub("[ +\\-]", "_", metadata$cell_type)
metadata$samples <- gsub("[ +\\-]", "_", metadata$samples)


# Update the metadata in the Seurat object
int_seurat_object@meta.data <- metadata

# Verify the changes
head(int_seurat_object@meta.data$cell_type)

print(unique(int_seurat_object$cell_type))

```



# Load NicheNet's ligand-receptor network and ligand-target matrix

```{r}
organism = "human"

options(timeout = 120)



if(organism == "human"){
  
  lr_network_all = 
    readRDS("lr_network_human_allInfo_30112033.rds") %>% 
    mutate(
      ligand = convert_alias_to_symbols(ligand, organism = organism), 
      receptor = convert_alias_to_symbols(receptor, organism = organism))
  
  lr_network_all = lr_network_all  %>% 
    mutate(ligand = make.names(ligand), receptor = make.names(receptor)) 
  
  lr_network = lr_network_all %>% 
    distinct(ligand, receptor)
  
  ligand_target_matrix = readRDS("ligand_target_matrix_nsga2r_final.rds")
  
  colnames(ligand_target_matrix) = colnames(ligand_target_matrix) %>% 
    convert_alias_to_symbols(organism = organism) %>% make.names()
  rownames(ligand_target_matrix) = rownames(ligand_target_matrix) %>% 
    convert_alias_to_symbols(organism = organism) %>% make.names()
  
  lr_network = lr_network %>% filter(ligand %in% colnames(ligand_target_matrix))
  ligand_target_matrix = ligand_target_matrix[, lr_network$ligand %>% unique()]
  
} else if(organism == "mouse"){
  
  lr_network_all = readRDS(url(
    "https://zenodo.org/record/10229222/files/lr_network_mouse_allInfo_30112033.rds"
    )) %>% 
    mutate(
      ligand = convert_alias_to_symbols(ligand, organism = organism), 
      receptor = convert_alias_to_symbols(receptor, organism = organism))
  
  lr_network_all = lr_network_all  %>% 
    mutate(ligand = make.names(ligand), receptor = make.names(receptor)) 
  lr_network = lr_network_all %>% 
    distinct(ligand, receptor)
  
  ligand_target_matrix = readRDS(url(
    "https://zenodo.org/record/7074291/files/ligand_target_matrix_nsga2r_final_mouse.rds"
    ))
  
  colnames(ligand_target_matrix) = colnames(ligand_target_matrix) %>% 
    convert_alias_to_symbols(organism = organism) %>% make.names()
  rownames(ligand_target_matrix) = rownames(ligand_target_matrix) %>% 
    convert_alias_to_symbols(organism = organism) %>% make.names()
  
  lr_network = lr_network %>% filter(ligand %in% colnames(ligand_target_matrix))
  ligand_target_matrix = ligand_target_matrix[, lr_network$ligand %>% unique()]
  
}

```
# Read in SingleCellExperiment Object

```{r}
sce = Seurat::as.SingleCellExperiment(int_seurat_object, assay = "RNA")
sce = alias_to_symbol_SCE(sce, "human") %>% makenames_SCE()

```

# Prepare the settings of the MultiNicheNet cell-cell communication analysis


## Define in which metadata columns we can find the group, sample and cell type IDs
```{r}
cell_metadata <- as.data.frame(colData(sce))

sample_id = "samples"
group_id = "timepoint"
celltype_id = "cell_type"
covariates = NA
batches = NA
```

## Define the contrasts of interest.
```{r}
# Just one contrast: Week_25 vs Week_38
contrasts_oi <- c("'Week_38-Week_25'")

contrast_tbl <- tibble(
  contrast = c("Week_38-Week_25"),
  group    = c("Week_25","Week_38")  # optional, or could label each explicitly
)
```

#Define the sender and receiver cell types of interest.

```{r}
senders_oi = SummarizedExperiment::colData(sce)[,celltype_id] %>% unique()
receivers_oi = SummarizedExperiment::colData(sce)[,celltype_id] %>% unique()
sce = sce[, SummarizedExperiment::colData(sce)[,celltype_id] %in% 
            c(senders_oi, receivers_oi)
          ]


```

# Cell-type filtering: determine which cell types are sufficiently present
```{r}
min_cells = 10

abundance_info = get_abundance_info(
  sce = sce, 
  sample_id = sample_id, group_id = group_id, celltype_id = celltype_id, 
  min_cells = min_cells, 
  )
gc()
```

```{r}
abundance_info$abund_plot_sample

```

# Cell type filtering based on cell type abundance information

```{r}
sample_group_celltype_df = abundance_info$abundance_data %>% 
  filter(n > min_cells) %>% 
  ungroup() %>% 
  distinct(sample_id, group_id) %>% 
  cross_join(
    abundance_info$abundance_data %>% 
      ungroup() %>% 
      distinct(celltype_id)
    ) %>% 
  arrange(sample_id)

abundance_df = sample_group_celltype_df %>% left_join(
  abundance_info$abundance_data %>% ungroup()
  )

abundance_df$n[is.na(abundance_df$n)] = 0
abundance_df$keep[is.na(abundance_df$keep)] = FALSE
abundance_df_summarized = abundance_df %>% 
  mutate(keep = as.logical(keep)) %>% 
  group_by(group_id, celltype_id) %>% 
  summarise(samples_present = sum((keep)))

celltypes_absent_one_condition = abundance_df_summarized %>% 
  filter(samples_present == 0) %>% pull(celltype_id) %>% unique() 
# find truly condition-specific cell types by searching for cell types 
# truely absent in at least one condition

celltypes_present_one_condition = abundance_df_summarized %>% 
  filter(samples_present >= 2) %>% pull(celltype_id) %>% unique() 
# require presence in at least 2 samples of one group so 
# it is really present in at least one condition

condition_specific_celltypes = intersect(
  celltypes_absent_one_condition, 
  celltypes_present_one_condition)

total_nr_conditions = SummarizedExperiment::colData(sce)[,group_id] %>% 
  unique() %>% length() 


absent_celltypes <- abundance_df_summarized %>%
  filter(samples_present < 2) %>%                # Keep rows where 'samples_present' is less than 2
  group_by(celltype_id) %>%                      # Group by 'celltype_id'
  summarize(n_conditions = n_distinct(group_id)) %>% # Count unique conditions per 'celltype_id'
  filter(n_conditions == total_nr_conditions) %>%    # Filter where count matches total conditions
  pull(celltype_id)    
  
print("condition-specific celltypes:")
## [1] "condition-specific celltypes:"
print(condition_specific_celltypes)
## character(0)
  
print("absent celltypes:")
## [1] "absent celltypes:"
print(absent_celltypes)
## character(0)
```

```{r}
analyse_condition_specific_celltypes = FALSE

if(analyse_condition_specific_celltypes == TRUE){
  senders_oi = senders_oi %>% setdiff(absent_celltypes)
  receivers_oi = receivers_oi %>% setdiff(absent_celltypes)
} else {
  senders_oi = senders_oi %>% 
    setdiff(union(absent_celltypes, condition_specific_celltypes))
  receivers_oi = receivers_oi %>% 
    setdiff(union(absent_celltypes, condition_specific_celltypes))
}

sce = sce[, SummarizedExperiment::colData(sce)[,celltype_id] %in% 
            c(senders_oi, receivers_oi)
          ]
gc()
```

# Gene filtering: determine which genes are sufficiently expressed in each present cell type
```{r}
min_sample_prop = 0.50
fraction_cutoff = 0.05

frq_list = get_frac_exprs(
  sce = sce, 
  sample_id = sample_id, celltype_id =  celltype_id, group_id = group_id, 
  batches = batches, 
  min_cells = min_cells, 
  fraction_cutoff = fraction_cutoff, min_sample_prop = min_sample_prop)

```


```{r}
genes_oi = frq_list$expressed_df %>% 
  filter(expressed == TRUE) %>% pull(gene) %>% unique() 
sce = sce[genes_oi, ]
gc()
```


# Pseudobulk expression calculation: determine and normalize per-sample pseudobulk expression levels for each expressed gene in each present cell type


```{r}
abundance_expression_info = process_abundance_expression_info(
  sce = sce, 
  sample_id = sample_id, group_id = group_id, celltype_id = celltype_id, 
  min_cells = min_cells, 
  senders_oi = senders_oi, receivers_oi = receivers_oi, 
  lr_network = lr_network, 
  batches = batches, 
  frq_list = frq_list, 
  abundance_info = abundance_info)
```

```{r}
gc()
```

```{r}
abundance_expression_info$celltype_info$pb_df %>% head()
abundance_expression_info$celltype_info$pb_df_group %>% head()
abundance_expression_info$sender_receiver_info$pb_df %>% head()
```


# Differential expression (DE) analysis: determine which genes are differentially expressed
```{r}
DE_info = get_DE_info(
  sce = sce, 
  sample_id = sample_id, group_id = group_id, celltype_id = celltype_id, 
  batches = batches, covariates = covariates, 
  contrasts_oi = contrasts_oi, 
  min_cells = min_cells, 
  expressed_df = frq_list$expressed_df)

```

```{r}
DE_info$hist_pvals
```
```{r}
empirical_pval = FALSE

if(empirical_pval == TRUE){
  DE_info_emp = get_empirical_pvals(DE_info$celltype_de$de_output_tidy)
  celltype_de = DE_info_emp$de_output_tidy_emp %>% select(-p_val, -p_adj) %>% 
    rename(p_val = p_emp, p_adj = p_adj_emp)
} else {
  celltype_de = DE_info$celltype_de$de_output_tidy
} 


```

# Combine DE information for ligand-senders and receptors-receivers
```{r}
sender_receiver_de = combine_sender_receiver_de(
  sender_de = celltype_de,
  receiver_de = celltype_de,
  senders_oi = senders_oi,
  receivers_oi = receivers_oi,
  lr_network = lr_network
)
```
 
# Combine DE information for ligand-senders and receptors-receivers
```{r}
sender_receiver_de = combine_sender_receiver_de(
  sender_de = celltype_de,
  receiver_de = celltype_de,
  senders_oi = senders_oi,
  receivers_oi = receivers_oi,
  lr_network = lr_network
)
```

```{r}
sender_receiver_de %>% head(20)
```
# Ligand activity prediction: use the DE analysis output to predict the activity of ligands in receiver cell types and infer their potential target genes

## Assess geneset_oi-vs-background ratios for different DE output tresholds prior to the NicheNet ligand activity analysis

```{r}
logFC_threshold = 0.5
p_val_threshold = 0.05
```
##
```{r}
p_val_adj = TRUE
geneset_assessment = contrast_tbl$contrast %>% 
  lapply(
    process_geneset_data, 
    celltype_de, logFC_threshold, p_val_adj, p_val_threshold
  ) %>% 
  bind_rows()
geneset_assessment
```
Use the adjusted p-value
```{r}
geneset_assessment_adjustedPval = contrast_tbl$contrast %>% 
  lapply(
    process_geneset_data, 
    celltype_de, logFC_threshold, p_val_adj = TRUE, p_val_threshold
    ) %>% 
  bind_rows() 
geneset_assessment_adjustedPval
```
## Perform the ligand activity analysis and ligand-target inference
```{r}
top_n_target = 250
```

```{r}
verbose = TRUE
cores_system = 7
n.cores = min(cores_system, celltype_de$cluster_id %>% unique() %>% length()) 
```

```{r}
rm(seurat_object)
rm(int_seurat_object)
rm(frq_list)
gc()
```

```{r}
ligand_activities_targets_DEgenes = suppressMessages(suppressWarnings(
  get_ligand_activities_targets_DEgenes(
    receiver_de = celltype_de,
    receivers_oi = intersect(receivers_oi, celltype_de$cluster_id %>% unique()),
    ligand_target_matrix = ligand_target_matrix,
    logFC_threshold = logFC_threshold,
    p_val_threshold = p_val_threshold,
    p_val_adj = p_val_adj,
    top_n_target = top_n_target,
    verbose = verbose, 
    n.cores = n.cores
  )
))
```

```{r}
ligand_activities_targets_DEgenes$ligand_activities %>% head(20)
```
# Prioritization: rank cell-cell communication patterns through multi-criteria prioritization
```{r}
ligand_activity_down = FALSE
```

```{r}
sender_receiver_tbl = sender_receiver_de %>% distinct(sender, receiver)

metadata_combined = SummarizedExperiment::colData(sce) %>% tibble::as_tibble()

if(!is.na(batches)){
  grouping_tbl = metadata_combined[,c(sample_id, group_id, batches)] %>% 
    tibble::as_tibble() %>% distinct()
  colnames(grouping_tbl) = c("sample","group",batches)
} else {
  grouping_tbl = metadata_combined[,c(sample_id, group_id)] %>% 
    tibble::as_tibble() %>% distinct()
  colnames(grouping_tbl) = c("sample","group")
}

prioritization_tables = suppressMessages(generate_prioritization_tables(
    sender_receiver_info = abundance_expression_info$sender_receiver_info,
    sender_receiver_de = sender_receiver_de,
    ligand_activities_targets_DEgenes = ligand_activities_targets_DEgenes,
    contrast_tbl = contrast_tbl,
    sender_receiver_tbl = sender_receiver_tbl,
    grouping_tbl = grouping_tbl,
    scenario = "regular", # all prioritization criteria will be weighted equally
    fraction_cutoff = fraction_cutoff, 
    abundance_data_receiver = abundance_expression_info$abundance_data_receiver,
    abundance_data_sender = abundance_expression_info$abundance_data_sender,
    ligand_activity_down = ligand_activity_down
  ))
```

# Calculate the across-samples expression correlation between ligand-receptor pairs and target genes 
```{r}
lr_target_prior_cor = lr_target_prior_cor_inference(
  receivers_oi = prioritization_tables$group_prioritization_tbl$receiver %>% unique(), 
  abundance_expression_info = abundance_expression_info, 
  celltype_de = celltype_de, 
  grouping_tbl = grouping_tbl, 
  prioritization_tables = prioritization_tables, 
  ligand_target_matrix = ligand_target_matrix, 
  logFC_threshold = logFC_threshold, 
  p_val_threshold = p_val_threshold, 
  p_val_adj = p_val_adj
  )

```

# Save all the output of MultiNicheNet
```{r}
path = "./"

multinichenet_output = list(
    celltype_info = abundance_expression_info$celltype_info,
    celltype_de = celltype_de,
    sender_receiver_info = abundance_expression_info$sender_receiver_info,
    sender_receiver_de =  sender_receiver_de,
    ligand_activities_targets_DEgenes = ligand_activities_targets_DEgenes,
    prioritization_tables = prioritization_tables,
    grouping_tbl = grouping_tbl,
    lr_target_prior_cor = lr_target_prior_cor
  ) 
multinichenet_output = make_lite_output(multinichenet_output)

save = TRUE
if(save == TRUE){
  saveRDS(multinichenet_output, paste0(path, "all_samples_multinichenet_output_v3.rds"))

}
```


