---
title: "scRNA of in vitro human skeletal stem cells"
author: "Lukas Moser"
date: today
date-format: long
format:
  html:
    toc: true
    self-contained: true
---

## Importing file from Trailmaker and do Datacleaning
```{r,include=FALSE}
set.seed(123)
#Read the input
library(openxlsx)
library(Seurat)
library(ggplot2)
library(cowplot)
library(dplyr)
library("Seurat")
library("limma")
library("clustree")
library("viridis")
library("glue")
library("knitr")
library("BiocParallel")
library("here")
library("writexl")
library("jsonlite")
library("tidyverse")
seurat_object <- readRDS("6be4e6e0-90bc-49ae-91ca-7a428ec7365c_processed_matrix.rds")
```

### Create a new Seurat object

```{r}
old_seurat_obj <- seurat_object

old_seurat_obj@assays$RNA@counts@Dimnames[[1]] <- old_seurat_obj@misc$gene_annotations$original_name

# Step 1: Extract counts data
counts_matrix <- old_seurat_obj[["RNA"]]@counts
# Step 2: Extract cell metadata
cell_metadata <- old_seurat_obj@meta.data
# Step 3: Create a new Seurat object
new_seurat_obj <- CreateSeuratObject(counts = counts_matrix, meta.data = cell_metadata)

old_seurat_obj <- seurat_object
seurat_object <- new_seurat_obj
rm(new_seurat_obj)
rm(old_seurat_obj)
rm(counts_matrix)
```

### QC with Mitochondria genes and doublets
```{r}
metadata <- seurat_object [[]]
metadata_first_10 <- head(seurat_object@meta.data, 10)
```

```{r}
library(dplyr)

# Create a summary table grouped by sample
sample_summary <- seurat_object@meta.data %>%
  group_by(samples) %>%
  summarise(
    Total_Cells = n(),
    Singlet_Cells = sum(doublet_class == "singlet", na.rm = TRUE),
    Percent_Singlet = (Singlet_Cells / Total_Cells) * 100
  )

# Display the summary
print(sample_summary)
```


```{r}
# Visualize the percent.mt and doublet_score columns as histograms
ggplot(seurat_object@meta.data, aes(x = percent.mt)) +
  geom_histogram(binwidth = 0.01, fill = 'blue', alpha = 0.7) +
  ggtitle('Histogram of percent.mt') +
  xlab('percent.mt') +
  ylab('Frequency')

ggplot(seurat_object@meta.data, aes(x = doublet_scores)) +
  geom_histogram(binwidth = 0.01, fill = 'red', alpha = 0.7) +
  ggtitle('Histogram of doublet_score') +
  xlab('doublet_score') +
  ylab('Frequency')

# Filter cells with percent.mt > 0.5 or doublet_score > 0.5
filtered_seurat_object <- subset(seurat_object, subset = percent.mt <= 1 & doublet_class <= "singlet")

# Calculate the number of cells removed
num_cells_removed <- ncol(seurat_object) - ncol(filtered_seurat_object)

# Print the number of cells removed
print(paste("Number of cells removed:", num_cells_removed))

# Return the filtered Seurat object
filtered_seurat_object
seurat_object <- filtered_seurat_object
```
### Normalize Data and find Variables Gene
```{r}
seurat_object <- NormalizeData(seurat_object, normalization.method = "LogNormalize", scale.factor = 10000)
seurat_object <- FindVariableFeatures(seurat_object, selection.method = "vst", nfeatures = 2000)
top10 <- head(VariableFeatures(seurat_object), 10)
```

### Run PCA and UMAP
```{r}
#invisible(ulimit::memory_limit(18000))
#I use here only the varibale feature instead of all, otherwise I run into memory issues
seurat_object <- ScaleData(seurat_object)
seurat_object <- RunPCA(seurat_object, features = VariableFeatures(object = seurat_object))

```
```{r}
DimHeatmap(seurat_object, dims = 1:20, cells = 500, balanced = TRUE)
ElbowPlot(seurat_object)
```
```{r}
seurat_object <- RunUMAP(seurat_object, dims = 1:15)
```
### Make Sample Groups
`````{r}

# Extract the metadata

metadata <- seurat_object@meta.data

# Create a new column by removing the last underscore and the number afterwards
metadata$group <- sub("_[^_]+$", "", metadata$samples)

# Assign the updated metadata back to the Seurat object
seurat_object@meta.data <- metadata

# Verify the changes
head(metadata)

# Optional: Print unique groups to check
unique_groups <- unique(metadata$group)
print(unique_groups)
```


```{r}
# Load necessary library
library(dplyr)

# Extract the metadata
metadata <- seurat_object@meta.data

# Create a summary table for the number of cells in each group
group_summary <- metadata %>%
  dplyr::group_by(group) %>%
  dplyr::summarize(total_cells = n()) %>%
  dplyr::arrange(desc(total_cells))

# Print the summary table for groups
print("Group summary:")
print(group_summary)

# Create a summary table for the number of cells in each sample within each group
sample_summary <- metadata %>%
  dplyr::group_by(group, samples) %>%
  dplyr::summarize(total_cells = n()) %>%
  dplyr::arrange(group, desc(total_cells))

# Print the summary table for samples within each group
print("Sample summary:")
print(sample_summary)

```

### Check UMAP
```{r}
Idents(seurat_object) <- "group"
DimPlot(seurat_object, reduction = "umap")
```

```{r}
#test <- seurat_object@misc$gene_annotations
gc()
```
# Assign Celltype


### Remove cultured cells (for another project used)
```{r}
# Define the groups to keep
groups_to_keep <- c("Preterm_Amnion", "Term_Amnion", "MMC_Amnion", "Placenta_Amnion",
                    "Preterm_Chorion", "Term_Chorion", "MMC_Chorion", "Placenta_Chorion")
# Extract the metadata
metadata <- seurat_object@meta.data
# Filter the metadata to keep only the cells in the specified groups
cells_to_keep <- rownames(metadata[metadata$group %in% groups_to_keep, ])
# Subset the Seurat object to keep only the selected cells
amnion_seurat_object <- subset(seurat_object, cells = cells_to_keep)

#rm(seurat_object)
gc()
```
### Preprocess Data

```{r}
amnion_seurat_object <- NormalizeData(amnion_seurat_object, normalization.method = "LogNormalize", scale.factor = 10000)
amnion_seurat_object <- FindVariableFeatures(amnion_seurat_object, selection.method = "vst", nfeatures = 3000)
#Again just run the the most varibale to scale the data
amnion_seurat_object <- ScaleData(amnion_seurat_object)

s.genes <- cc.genes$s.genes
g2m.genes <- cc.genes$g2m.genes
amnion_seurat_object <- CellCycleScoring(amnion_seurat_object, s.features = s.genes, g2m.features = g2m.genes, set.ident = TRUE)
# all.genes <- rownames(amnion_seurat_object)
# amnion_seurat_object <- ScaleData(amnion_seurat_object, features = all.genes)
amnion_seurat_object <- RunPCA(amnion_seurat_object, features = VariableFeatures(object = amnion_seurat_object))

gc()
#amnion_seurat_object<- RunPCA(amnion_seurat_object, features = VariableFeatures(object = amnion_seurat_object))
#VizDimLoadings(amnion_seurat_object, dims = 1:2, reduction = "pca")
DimHeatmap(amnion_seurat_object, dims = 1:21, cells = 500, balanced = TRUE)
ElbowPlot(amnion_seurat_object)


amnion_seurat_object <- RunTSNE(amnion_seurat_object, dims = 1:20)
amnion_seurat_object <- RunUMAP(amnion_seurat_object, dims = 1:20)
Idents(amnion_seurat_object) <- "group"
DimPlot(amnion_seurat_object, reduction = "umap")

```

###Clustering

#### Clustering Algortihm
```{r}
metadata <- amnion_seurat_object@meta.data
n.dims <- 20
resolutions <- seq(0.2, 1.8, 0.2)
amnion_seurat_object <- FindNeighbors(amnion_seurat_object, dims = 1:20)
amnion_seurat_object <- FindClusters(amnion_seurat_object, reduction.type = "cca.aligned",
                         dims.use = 1:n.dims, resolution = resolutions)
metadata <- amnion_seurat_object@meta.data

```


#### Umap of the Clustering
```{r}
# Define the resolutions
#resolutions <- c(0.6, 0.8, 1.0, 1.2, 1.4)

# Initialize lists to store UMAP, t-SNE plots and top markers for each resolution
umap_plots <- list()
tsne_plots <- list()
  
# Loop through each resolution
for (res in resolutions) {
  # Set the identity of the Seurat object to the respective resolution
  Idents(amnion_seurat_object) <- paste0("RNA_snn_res.", res)
  
  
  # Generate UMAP and t-SNE plots
  umap_plot <- DimPlot(amnion_seurat_object, reduction = "umap", group.by = paste0("RNA_snn_res.", res)) + ggtitle(paste("UMAP - Res:", res))
  tsne_plot <- DimPlot(amnion_seurat_object, reduction = "tsne", group.by = paste0("RNA_snn_res.", res)) + ggtitle(paste("t-SNE - Res:", res))
  
  # Save plots to lists
  umap_plots[[paste0("res_", res)]] <- umap_plot
  tsne_plots[[paste0("res_", res)]] <- tsne_plot
  
}

# Combine UMAP and t-SNE plots for each resolution
combined_plots <- list()
for (res in resolutions) {
  combined_plot <- umap_plots[[paste0("res_", res)]] | tsne_plots[[paste0("res_", res)]]
  combined_plots[[paste0("res_", res)]] <- combined_plot
  print(combined_plot)
}
```
#### Clustertree
```{r}
#From here, other interesting plots also here: http://oshlacklab.com/combes-organoid-paper/07_Combined_Clustering.html#
clustree(amnion_seurat_object)
```

```{r}
clustree(amnion_seurat_object, node_colour = "sc3_stability")
```


# Assign Celltype according to Pique-Regi


#### Based on Pique-Regi Paper with  scType

```{r}
library(HGNChelper)
library(openxlsx)
source("https://raw.githubusercontent.com/IanevskiAleksandr/sc-type/master/R/gene_sets_prepare.R"); source("https://raw.githubusercontent.com/IanevskiAleksandr/sc-type/master/R/sctype_score_.R")
# load gene set preparation function
source("https://raw.githubusercontent.com/IanevskiAleksandr/sc-type/master/R/gene_sets_prepare.R")
# load cell type annotation function
source("https://raw.githubusercontent.com/IanevskiAleksandr/sc-type/master/R/sctype_score_.R")
```

```{r}
# Load necessary libraries
library(readxl)
library(dplyr)
library(writexl)

# File and gene setup
marker_file <- "Pique_et_all_2019_markergenes.xls"
tissue_type <- "amnion"
synaptic_neuron_gene <- ""  # Replace with the actual gene name you want

# Read the Excel file
markers_df <- read_excel(marker_file)

# Summarize the DataFrame
final_df <- markers_df %>%
  group_by(cluster) %>%
  summarise(
    geneSymbolmore1 = paste(gene, collapse = ", "),
    .groups = 'drop'
  ) %>%
  mutate(
    tissueType = tissue_type,
    cellName = cluster,
    geneSymbolmore2 = synaptic_neuron_gene  # Set specific gene for synaptic neurons
  ) %>%
  dplyr::select(tissueType, cellName, geneSymbolmore1, geneSymbolmore2)

# Write to an Excel file
output_file <- "gs.xlsx"
write_xlsx(final_df, path = output_file)

# Call the gene_sets_prepare function (make sure it's loaded or available)
gs_list <- gene_sets_prepare(output_file, tissue_type)


```



```{r}

# check Seurat object version (scRNA-seq matrix extracted differently in Seurat v4/v5)
seurat_package_v5 <- isFALSE('counts' %in% names(attributes(amnion_seurat_object[["RNA"]])));
print(sprintf("Seurat object %s is used", ifelse(seurat_package_v5, "v5", "v4")))

# extract scaled scRNA-seq matrix
scRNAseqData_scaled <- if (seurat_package_v5) as.matrix(amnion_seurat_object[["RNA"]]$scale.data) else as.matrix(amnion_seurat_object[["RNA"]]@scale.data)

# run ScType
es.max <- sctype_score(scRNAseqData = scRNAseqData_scaled, scaled = TRUE, gs = gs_list$gs_positive, gs2 = gs_list$gs_negative)

# NOTE: scRNAseqData parameter should correspond to your input scRNA-seq matrix. For raw (unscaled) count matrix set scaled = FALSE
# When using Seurat, we use "RNA" slot with 'scale.data' by default. Please change "RNA" to "SCT" for sctransform-normalized data,
# or to "integrated" for joint dataset analysis. To apply sctype with unscaled data, use e.g. amnion_seurat_object[["RNA"]]$counts or amnion_seurat_object[["RNA"]]@counts, with scaled set to FALSE.

# merge by cluster
cL_resutls <- do.call("rbind", lapply(unique(amnion_seurat_object@meta.data$RNA_snn_res.1), function(cl){
    es.max.cl = sort(rowSums(es.max[ ,rownames(amnion_seurat_object@meta.data[amnion_seurat_object@meta.data$RNA_snn_res.1==cl, ])]), decreasing = !0)
    head(data.frame(cluster = cl, type = names(es.max.cl), scores = es.max.cl, ncells = sum(amnion_seurat_object@meta.data$RNA_snn_res.1==cl)), 10)
}))
sctype_scores <- cL_resutls %>% group_by(cluster) %>% top_n(n = 1, wt = scores)  

# set low-confident (low ScType score) clusters to "unknown"
sctype_scores$type[as.numeric(as.character(sctype_scores$scores)) < sctype_scores$ncells/4] <- "Unknown"
print(sctype_scores[,1:3])

```
```{r}
Idents(amnion_seurat_object) <- "RNA_snn_res.1"

# Initialize a new metadata column for cell types
amnion_seurat_object$scType_Pique <- NA

# Loop through each cluster in sctype_scores and assign the corresponding type
for (i in 1:nrow(sctype_scores)) {
  cluster_id <- as.character(sctype_scores$cluster[i])
  cell_type <- sctype_scores$type[i]
  
  # Get cells belonging to the current cluster
  cells_in_cluster <- WhichCells(amnion_seurat_object, idents = cluster_id)
  
  # Assign the cell type to these cells
  amnion_seurat_object$scType_Pique[cells_in_cluster] <- cell_type
}

```



```{r}
DimPlot(amnion_seurat_object, reduction = "umap", group.by = "scType_Pique", label = TRUE,
  label.size = 3)+ NoLegend()
DimPlot(amnion_seurat_object, reduction = "tsne", group.by = "scType_Pique", label = TRUE,
  label.size = 3)+ NoLegend()

```

#### Make the unkown cluster to epithelial cells

```{r}
# Change all 'Unknown' entries in the 'scType_Pique' column to 'Epithelial'
amnion_seurat_object$scType_Pique[amnion_seurat_object$scType_Pique == "Unknown"] <- "Epithelial"
DimPlot(amnion_seurat_object, reduction = "umap", group.by = "scType_Pique", label = TRUE,
  label.size = 3)+ NoLegend()
```
# Find the origin

# Classify Sex of cells
```{r}
library(speckle)
library(SingleCellExperiment)
library(cellXY)
library(CellBench)
library(org.Hs.eg.db)

# Get the total number of cells in your Seurat object
total_cells <- ncol(amnion_seurat_object)

# Split cells into three groups
group1 <- sample(colnames(amnion_seurat_object), size = floor(total_cells / 3))
remaining_cells <- setdiff(colnames(amnion_seurat_object), group1)
group2 <- sample(remaining_cells, size = floor(total_cells / 3))
group3 <- setdiff(remaining_cells, group2)

# Function to classify sex for a subset of cells
classify_group_sex <- function(cells_subset) {
    # Subset the Seurat object
    seurat_subset <- subset(amnion_seurat_object, cells = cells_subset)
    
    # Extract the counts matrix
    counts_matrix <- GetAssayData(seurat_subset, layer = "counts", assay = "RNA")
    
    # Run classifySex
    sex <- classifySex(counts_matrix, genome = "Hs")
    
    return(sex)
}

# Run classifySex for each group
sex_group1 <- classify_group_sex(group1)
sex_group2 <- classify_group_sex(group2)
sex_group3 <- classify_group_sex(group3)

# Combine sex_group1 and sex_group2 by rows
combined_sex <- rbind(sex_group1, sex_group2)

# Now append sex_group3 to the combined data frame
combined_sex <- rbind(combined_sex, sex_group3)

metadata <- amnion_seurat_object@meta.data

rownames(metadata) <- rownames(metadata)
rownames(combined_sex) <- rownames(combined_sex)

combined_data <- merge(metadata, combined_sex, by = "row.names", all = TRUE)

rownames(combined_data) <- combined_data$Row.names
combined_data$Row.names <- NULL

amnion_seurat_object@meta.data <- combined_data
```

```{r}
# Your donor mapping
donor_df <- data.frame(
  samples = c(
    "Term_Amnion_1", "Term_Amnion_2", "Term_Amnion_3", "Term_Amnion_4", "Term_Amnion_5",
    "Term_Chorion_1", "Term_Chorion_2", "Term_Chorion_3", "Term_Chorion_4", "Term_Chorion_5",
    "Preterm_Amnion_1", "Preterm_Amnion_2", "Preterm_Amnion_3", "Preterm_Amnion_4",
    "Preterm_Chorion_1", "Preterm_Chorion_2", "Preterm_Chorion_3", "Preterm_Chorion_4",
    "MMC_Amnion_1000", "MMC_Amnion_1001", "MMC_Amnion_1002", "MMC_Amnion_1003", "MMC_Amnion_1004", "MMC_Amnion_1005",
    "MMC_Chorion_1000", "MMC_Chorion_1001", "MMC_Chorion_1002", "MMC_Chorion_1003", "MMC_Chorion_1004", "MMC_Chorion_1005",
    "Placenta_Amnion_3","Placenta_Chorion_3","Placenta_Amnion_1","Placenta_Chorion_1","Placenta_Amnion_2","Placenta_Chorion_2"
  ),
  Donor_Nr = c(
    51, 48, 45, 46, 49,
    51, 48, 45, 46, 49,
    53, 26, 41, 10,
    53, 26, 41, 10,
    1000, 1001, 1002, 1003, 1004, 1005,
    1000, 1001, 1002, 1003, 1004, 1005,
    51,51,45,45,46,46
  ))

# Create named vector: sample names as names, donor numbers as values
donor_map <- setNames(donor_df$Donor_Nr, donor_df$samples)

amnion_seurat_object@meta.data$Donor_Nr <- 1
# Apply mapping directly to create a new Donor_Nr column in Seurat metadata
amnion_seurat_object@meta.data$Donor_Nr <- donor_map[amnion_seurat_object@meta.data$samples]

metadata <- amnion_seurat_object@meta.data
```

```{r}

# 1. Create the folder if it doesn't exist
output_dir <- "SI_Sexassignment"
if (!dir.exists(output_dir)) {
  dir.create(output_dir)
}

# 2. Define the groups and filter cells
groups_to_keep <- c(45, 49, 26, 41, 10, 1005)
metadata <- amnion_seurat_object@meta.data
cells_to_keep <- rownames(metadata[metadata$Donor_Nr %in% groups_to_keep, ])

# 3. Generate and save Plot 1: Prediction UMAP
p1 <- DimPlot(amnion_seurat_object, reduction = "umap", cells = cells_to_keep, group.by = "prediction")
ggsave(filename = file.path(output_dir, "umap_prediction.png"), plot = p1, width = 7, height = 6, dpi = 300)
print(p1)

# 4. Generate and save Plot 2: scType_Pique UMAP
p2 <- DimPlot(amnion_seurat_object, reduction = "umap", cells = cells_to_keep, group.by = "scType_Pique", label = TRUE, label.size = 3) + NoLegend()
ggsave(filename = file.path(output_dir, "umap_scType_Pique.png"), plot = p2, width = 7, height = 6, dpi = 300)
print(p2)

# 5. Generate and save Plot 3: XIST FeaturePlot
p3 <- FeaturePlot(amnion_seurat_object, "XIST", reduction = "umap", cells = cells_to_keep)
ggsave(filename = file.path(output_dir, "featureplot_XIST.png"), plot = p3, width = 7, height = 6, dpi = 300)
print(p3)
```

```{r}
male_fetus <- c(45,49,26,41,10,1005)
amnion_no_graphs <- amnion_seurat_object
amnion_no_graphs@graphs <- list()
filtered_seurat_object <- subset(amnion_no_graphs, subset = Donor_Nr %in% male_fetus)
metadata <- filtered_seurat_object@meta.data
# Load required libraries
library(dplyr)
library(ggplot2)
library(tidyr)

# Assuming metadata is already defined
# Ensure `cell_type` and `scType_Pique` are characters
metadata <- metadata %>%
  mutate(cell_type = as.character(cell_type), 
         scType_Pique = as.character(scType_Pique))

# Calculate the ratio of male and female cells for each cluster
ratio_df <- metadata %>%
  filter(!is.na(prediction)) %>%
  group_by(RNA_snn_res.1, prediction) %>%
  summarise(count = n(), .groups = 'drop') %>%
  pivot_wider(names_from = prediction, values_from = count, values_fill = 0) %>%
  mutate(total = Male + Female,
         Male_ratio = Male / total)

# Get cluster information for titles
cluster_info <- metadata %>%
  group_by(RNA_snn_res.1) %>%
  summarise(cell_type = unique(cell_type)[1],  # Get the first unique cell_type
            scType_Pique = unique(scType_Pique)[1],  # Get the first unique scType_Pique
            .groups = 'drop')

# Join the ratio data with cluster info
ratio_df <- ratio_df %>%
  left_join(cluster_info, by = "RNA_snn_res.1")

# Convert RNA_snn_res.1 to factor to maintain order
ratio_df$RNA_snn_res.1 <- factor(ratio_df$RNA_snn_res.1, levels = unique(ratio_df$RNA_snn_res.1))

# Plot 3: Total Number of Cells by scType_Pique
ggplot(ratio_df, aes(x = RNA_snn_res.1, y = total, fill = "Total Cells")) +
  geom_bar(stat = "identity", fill = "red") +
  labs(x = "Clusters", 
       y = "Total Cells", 
       title = "Total Number of Cells by Cluster (scType_Pique)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) +
  scale_x_discrete(labels = ratio_df$scType_Pique)

# Plot 4: Male Ratio by scType_Pique
ggplot(ratio_df, aes(x = RNA_snn_res.1, y = Male_ratio, fill = "Male Ratio")) +
  geom_bar(stat = "identity", fill = "blue") +
  labs(x = "Clusters", 
       y = "Male Ratio", 
       title = "Male Ratio by Cluster (scType_Pique)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) +
  scale_x_discrete(labels = ratio_df$scType_Pique)

amnion_ratio_df <- ratio_df
```

```{r}
library(dplyr)
library(ggplot2)
library(tidyr)
library(Seurat)

# 1. Define output directory and ensure it exists
output_dir <- "SI_Sexassignment"
if (!dir.exists(output_dir)) {
  dir.create(output_dir)
}

# 2. Data Subset & Preparation
male_fetus <- c(45, 49, 26, 41, 10, 1005)
amnion_no_graphs <- amnion_seurat_object
amnion_no_graphs@graphs <- list()

filtered_seurat_object <- subset(amnion_no_graphs, subset = Donor_Nr %in% male_fetus)
metadata <- filtered_seurat_object@meta.data

metadata <- metadata %>%
  mutate(cell_type = as.character(cell_type), 
         scType_Pique = as.character(scType_Pique))

# Calculate ratios
ratio_df <- metadata %>%
  filter(!is.na(prediction)) %>%
  group_by(RNA_snn_res.1, prediction) %>%
  summarise(count = n(), .groups = 'drop') %>%
  pivot_wider(names_from = prediction, values_from = count, values_fill = 0) %>%
  mutate(total = Male + Female,
         Male_ratio = Male / total)

# Get cluster information
cluster_info <- metadata %>%
  group_by(RNA_snn_res.1) %>%
  summarise(cell_type = unique(cell_type)[1], 
            scType_Pique = unique(scType_Pique)[1], 
            .groups = 'drop')

ratio_df <- ratio_df %>% left_join(cluster_info, by = "RNA_snn_res.1")
ratio_df$RNA_snn_res.1 <- factor(ratio_df$RNA_snn_res.1, levels = unique(ratio_df$RNA_snn_res.1))
amnion_ratio_df <- ratio_df

# ==================== GENERATING & SAVING PLOTS ====================

# Plot A: Male Ratio by scType_Pique (Your original plot)
p_ratio_pique <- ggplot(ratio_df, aes(x = RNA_snn_res.1, y = Male_ratio)) +
  geom_bar(stat = "identity", fill = "blue") +
  labs(x = "Cell Types (Pique)", 
       y = "Male Ratio", 
       title = "Male Ratio by Cluster (scType_Pique)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) +
  scale_x_discrete(labels = ratio_df$scType_Pique)

ggsave(filename = file.path(output_dir, "male_ratio_by_scType_Pique.png"), plot = p_ratio_pique, width = 8, height = 6, dpi = 300)


# Plot B: Male Ratio by Cluster Number (X-axis changed to cluster numbers)
p_ratio_cluster <- ggplot(ratio_df, aes(x = RNA_snn_res.1, y = Male_ratio)) +
  geom_bar(stat = "identity", fill = "blue") +
  labs(x = "Cluster ID", 
       y = "Male Ratio", 
       title = "Male Ratio by Cluster Number") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 0, hjust = 0.5)) # Kept flat since numbers are short

ggsave(filename = file.path(output_dir, "male_ratio_by_cluster_number.png"), plot = p_ratio_cluster, width = 8, height = 6, dpi = 300)


# Plot C: UMAP of Labeled Clusters (On the filtered subset)
p_umap_clusters <- DimPlot(filtered_seurat_object, 
                           reduction = "umap", 
                           group.by = "RNA_snn_res.1", 
                           label = TRUE, 
                           label.size = 4) + 
  labs(title = "UMAP of Filtered Subsets by Cluster ID") +
  NoLegend()

ggsave(filename = file.path(output_dir, "umap_filtered_clusters.png"), plot = p_umap_clusters, width = 7, height = 6, dpi = 300)

```



```{r}
amnion_ratio_df$origin <- ifelse(amnion_ratio_df$Male_ratio > 0.70, "Fetus", "Mother")

# Check that the cluster column in your Seurat metadata matches type with amnion_ratio_df
amnion_ratio_df$RNA_snn_res.1 <- as.character(amnion_ratio_df$RNA_snn_res.1)
amnion_seurat_object@meta.data$RNA_snn_res.1 <- as.character(amnion_seurat_object@meta.data$RNA_snn_res.1)

# Create a lookup vector for origin by cluster
origin_map <- setNames(amnion_ratio_df$origin, amnion_ratio_df$RNA_snn_res.1)

# Map origin to each cell in metadata based on RNA_snn_res.1 cluster
amnion_seurat_object@meta.data$origin <- origin_map[amnion_seurat_object@meta.data$RNA_snn_res.1]
metadata <- amnion_seurat_object@meta.data
```

# Amnion or Chorion tissue

```{r}
amnion_seurat_object@meta.data$layer <- ifelse(grepl("Chorion", metadata$samples), "Chorion",
                         ifelse(grepl("Amnion", metadata$samples), "Amnion", NA))

amnion_seurat_object@meta.data <- amnion_seurat_object@meta.data %>%
  mutate(
    tissue_source = if_else(str_detect(samples, "Placenta"), "Placenta", "Peripher")
  )

metadata <- amnion_seurat_object@meta.data
```

# Which timepoint

```{r}
amnion_seurat_object@meta.data <- amnion_seurat_object@meta.data %>%
  mutate(timepoint = case_when(
    Donor_Nr %in% c(51, 48, 45, 46, 49) ~ "Week_38",
    Donor_Nr == 53 ~ "Week_30",
    Donor_Nr %in% c(26, 41, 10) ~ "Week_34",
    Donor_Nr %in% c(1000, 1001, 1002, 1003, 1004, 1005) ~ "Week_25",
    TRUE ~ NA_character_  # For anything else
  ))


amnion_seurat_object$gestational_timepoint <- ifelse(
  amnion_seurat_object$timepoint == "Week_38", "Term",
  ifelse(amnion_seurat_object$timepoint == "Week_25", "Midterm", "Preterm")
)

amnion_seurat_object@meta.data <- amnion_seurat_object@meta.data %>%
  mutate(
    timepoint_plus = ifelse(grepl("Placenta", samples),
                            paste0(timepoint, "_Placenta"),
                            timepoint),
    gestational_timepoint_plus = ifelse(grepl("Placenta", samples),
                                       paste0(gestational_timepoint, "_Placenta"),
                                       gestational_timepoint)
  )

amnion_seurat_object@meta.data <- amnion_seurat_object@meta.data %>%
  mutate(
    layer_and_timepoint = paste(layer, timepoint_plus, sep = "_"),
    layer_and_gestation = paste(layer, gestational_timepoint_plus, sep = "_")
  )

amnion_seurat_object@meta.data <- amnion_seurat_object@meta.data %>%
  mutate(
    cell_type = paste(scType_Pique,origin, sep = "_"))


metadata <- amnion_seurat_object@meta.data

print(unique(metadata$layer_and_timepoint))
```

```{r}

amnion_safe <- DietSeurat(
  object = amnion_seurat_object,
  assays = DefaultAssay(amnion_seurat_object),  # keep counts/data
  dimreducs = NULL,   # remove all PCA/UMAP/other reductions
  graphs = NULL,      # remove SNN/NN graphs
  features = rownames(amnion_seurat_object)
)

# Optional: reset active identities
Idents(amnion_safe) <- NULL

saveRDS(amnion_safe, file = "amnion_seurat_object.rds")
```


#End
















